---
title: "HW7_answer"
author: "Guojing Wu"
date: "4/15/2019"
output:
    pdf_document:
    highlight: default
    number_sections: true
    citation_package:
    keep_tex: false
    fig_caption: true
    latex_engine: pdflatex
fontsize: 10pt
geometry: margin=1in
bibliography:
biblio-style:
header-includes:
- \usepackage{indentfirst}
- \usepackage{graphicx}
- \usepackage{geometry}
- \usepackage{subfigure}
- \usepackage{amsmath}
- \usepackage{listings}
- \usepackage{tikz}
- \usetikzlibrary{matrix}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "")
library(tidyverse)
library(grid)
library(gridExtra)
library(nlme)
data.poli <- read.csv("HW7-politeness_data.csv", sep = ",")
```

## Problem 

### 1) Exploratory analysis

```{r, dpi = 300}
plot.gender <- data.poli %>% 
  ggplot(aes(x = gender, y = frequency)) +
  geom_boxplot()

plot.atti <- data.poli %>% 
  ggplot(aes(x = attitude, y = frequency)) +
  geom_boxplot()

grid.arrange(plot.gender, plot.atti, ncol = 2)
```

The boxplot shows that there are differencies of pitch between male and female and between formal and informal

### 2) Fixed effect model with random intercept

```{r}
LMM1 <- lme(frequency ~ gender + attitude, data = data.poli, random = ~1 | subject, method = 'REML')
```

The covariance matrix for a subject $Y_{i}$ is
$$
cov(Y_{i}) = 
\begin{pmatrix}
  \sigma_{b}^{2} + \sigma^{2} & \sigma_{b}^{2} & ... & \sigma_{b}^{2} \\
  \sigma_{b}^{2} & \sigma_{b}^{2} + \sigma^{2} &... & \sigma_{b}^{2} \\
  ... &  &  &  \\
  \sigma_{b}^{2} & \sigma_{b}^{2} & ...  & \sigma_{b}^{2} + \sigma^{2} \\
\end{pmatrix}
$$

with $\sigma_{b}^{2}$ = `r VarCorr(LMM1)[1]`, $\sigma^{2}$ = `r VarCorr(LMM1)[2]`.

And the covariance matrix for REML is:

```{r}
vcov(LMM1) %>% knitr::kable()
```

The BLUPs for subject-specific intercept

```{r}
random.effects(LMM1) %>% knitr::kable()
```

The residuals are:

```{r, dpi = 300}
# residual
resid.mat <- matrix(as.numeric(data.poli$frequency - fitted(LMM1)), nrow = 14, ncol = 6)
colnames(resid.mat) = rownames(random.effects(LMM1))
resid.mat %>% knitr::kable()

plot(density(as.numeric(resid.mat)), xlab = "residual", main = "residual distribution")
```

### 3) Fixed effects model with intercepts for different subjects.

```{r}
# interaction term
LMM.1 <- lme(frequency ~ gender + attitude, data = data.poli, random = ~1 | subject, method = 'ML')
LMM.2 <- lme(frequency ~ gender * attitude, data = data.poli, random = ~1 | subject, method = 'ML')
res.inter <- anova(LMM.1, LMM.2)
```

For likelihood ratio test, we use `ML` instead of `REML`, and then doing ANOVA. The pvalue = `r round(res.inter[[length(res.inter)]][2], 4)`, so in this case, we fail to reject the null hypothesis and state that the interaction term is not significantly associated with pitch.

### 4) Fixed effect model with random intercept and random slope

Consider the model form is 
$$
\begin{split}
Y_{i,j} 
&= \beta_{1} + \beta_{2}\cdot gender_{i,j} + \beta_{3}\cdot attitude_{i,j} + b_{1,i} + b_{2,i}attitude_{i,j} + \epsilon_{i,j} \\
&= (\beta_{1} + b_{1,i}) + \beta_{2}\cdot gender_{i,j} + (\beta_{3} + b_{2,i})\cdot attitude_{i,j} + \epsilon_{i,j}
\end{split}
$$
where $b_{1,i} \sim N(0, g_{11})$, $b_{2,i} \sim N(0, g_{22})$, $cov(b_{1,i}, b_{2,i}) = g_{12}$, $\epsilon_{i,j} \sim N(0, \sigma^{2})$, and $(b_{1,i}, b_{2,i})^{T}$ and $\epsilon$ are assumed to be independent.

```{r}
LMM2 <- lme(frequency ~ gender + attitude, random = ~ 1 + attitude | subject, data = data.poli)
```

To compute the covariance structure, the $g_{11}, g_{12}, g_{22}, \sigma$ can be obtained from below:

```{r}
VarCorr(LMM2) %>% knitr::kable()
```

The fixed effect is:

```{r}
fixed.effects(LMM2) %>% knitr::kable()
```

The random effect is:

```{r}
random.effects(LMM2) %>% knitr::kable()
```

The BLUP for the first female subject in scenario 1 with polite attitude is `r fitted(LMM2)[1]`.

## Code
```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```
